### 2022-04-20 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e10565101f449c7cd01760eebd7624592f2c71d9 🚀

\[
v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right)\right|}{2\pi}\right]\,,
\]

where X is \(atan(cos(deg2rad(60.0)), sin(deg2rad(60.0))\), a constant
approximately equal to \(0.5235\).

@@

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness (Isaac et al., 2007) for each of the viruses in the
tree, then averaged these at the bat species level, projected these
values onto their geographic distributions, and averaged across every
bat found in a given pixel. As such, this can be thought of as a map of
the mean evolutionary distinctiveness of the known viral community
believed to be associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from Kreft \& Jetz (2010), who
proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map.

Here, we build on this idea with an entirely novel methodology. First,
we measure the phylogenetic distance between the different viruses in
the betacoronavirus tree by using the cophenetic function in `ape';
subsequently, we take a principal components analysis of that distance
matrix (readily interchangeable for NMDS in this case) to project the
viral tree into an n-dimensional space. We then take the first two
principal components and, as with the evolutionary distinctiveness
analysis, aggregated these to a mean host value and projected them using
a four-color bivariate map.

@@

@@

@@

>>>>

\[
v\times\left[1-\frac{\left|\text{atan}\left(\text{cos}(\text{rad}(h)), \text{sin}(\text{rad}(h))\right) - X\right|}{2\pi}\right]\,,
\]

where X is \(atan(cos(deg2rad(60.0)), sin(deg2rad(60.0)))\), a constant
approximately equal to \(0.5235\).

@@

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness (Isaac et al. 2007) for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

@@

Subsequently, we tested the hypothesis that the biogeography of bat
betacoronaviruses should track the biogeography of their hosts. To test
this idea, we loosely adapted a method from (Kreft and Jetz 2007, 2010),
who proposed a phylogenetic method for the delineation of animal
biogeographic regions. In their original method, a distance matrix -
where each row or column represents a geographic raster's grid cell, and
the dissimilarity values are the ``beta diversity similarity'' of their
community assemble - undergoes non-metric multidimensional scaling
(NMDS); the first two axes of the NMDS are projected geographically
using a four-color bivariate map. Here, we build on this idea with an
entirely novel methodology. First, we measure the phylogenetic distance
between the different viruses in the betacoronavirus tree by using the
cophenetic function in \texttt{ape} (Paradis and Schliep 2019);
subsequently, we take a principal components analysis of that distance
matrix (readily interchangeable for NMDS in this case) to project the
viral tree into an n-dimensional space. We then take the first two
principal components and, as with the evolutionary distinctiveness
analysis, aggregated these to a mean host value and projected them using
a four-color bivariate map.

@@

\leavevmode\hypertarget{ref-Isaac2007MamEdg}{}%
Isaac, N. J. B. et al. 2007. Mammals on the EDGE: Conservation
Priorities Based on Threat and Phylogeny. - PLOS ONE 2: e296.

@@

\leavevmode\hypertarget{ref-Kreft2007GloPat}{}%
Kreft, H. and Jetz, W. 2007. Global patterns and determinants of
vascular plant diversity. - Proceedings of the National Academy of
Sciences 104: 5925--5930.

\leavevmode\hypertarget{ref-Kreft2010FraDel}{}%
Kreft, H. and Jetz, W. 2010. A framework for delineating biogeographical
regions based on species distributions. - Journal of Biogeography 37:
2029--2053.

@@

\leavevmode\hypertarget{ref-Paradis2019ApeEnv}{}%
Paradis, E. and Schliep, K. 2019. Ape 5.0: An environment for modern
phylogenetics and evolutionary analyses in R. - Bioinformatics 35:
526--528.
